---
title: "HVTN 703/704 Sieve Analysis; Variable Importance Analysis"
output: word_document
date: '2023-03-13'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(pROC)
library(SuperLearner)
library(mgcv)
library(knitr)
library(cvAUC)

amp <- read.csv("/Volumes/trials/vaccine/p704/analysis/sieve/adata/amp_sieve_pooled_marks_final_v9.csv")
amp$gmt80ls_num <- amp$gmt80ls
amp$gmt80ls_num[amp$gmt80ls == ">100"] <- 100
amp$gmt80ls_num <- as.numeric(amp$gmt80ls_num)

amp$tx_bin <- ifelse(amp$tx_pool == "T1+T2", 1, 0)
amp703 <- subset(amp, protocol == "HVTN 703" & !is.na(parscore1.ls) & hiv1event == 1)
amp704 <- subset(amp, protocol == "HVTN 704" & !is.na(parscore1.ls) & hiv1event == 1)

# Treatment variables: tx (trichotomous treatment); tx_pool (dichotomous treatment)
# 703 scores: parscore1.ls, parscore2.ls, epitope.dist.b.ls, 
#             hdist.zspace.sites.binding.all.ls, gmt80ls
# 704 scores: parscore1.ls, parscore2.ls, epitope.dist.c.ls,
#             hdist.zspace.sites.binding.all.ls, gmt80ls
# 703 location: southAfrica
# 704 location: southAmerica

```

## Objective 1: Prediction accuracy of models to predict TX assignment using a single including all Env AA feature

```{r Objective1-703}
set.seed(300)
n <- nrow(amp703)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit703.parscore1.aucs <- numeric(5)

# Parscore 1
testfit703.parscore1 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.parscore1 <- mgcv::gam(tx_bin ~ s(parscore1.ls) + southAfrica,
                                 family = binomial, data = train.dat,
                                 method = "NCV")
  fitted703.parscore1 <- predict(trainfit703.parscore1, newdata = test.dat,
                                 type = "response")
  
  testfit703.parscore1[folds == j] <- fitted703.parscore1
}

result <- ci.cvAUC(labels = amp703$tx_bin, predictions = testfit703.parscore1, folds = folds)
fit703.parscore1.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Parscore 2
testfit703.parscore2 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.parscore2 <- mgcv::gam(tx_bin ~ s(parscore2.ls) + southAfrica,
                                     family = binomial, data = train.dat,
                                     method = "NCV")
  fitted703.parscore2 <- predict(trainfit703.parscore2, newdata = test.dat,
                                 type = "response")
  
  testfit703.parscore2[folds == j] <- fitted703.parscore2
}

result <- ci.cvAUC(labels = amp703$tx_bin, predictions = testfit703.parscore2, folds = folds)
fit703.parscore2.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Epitope dist
testfit703.epitope <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.epitope <- mgcv::gam(tx_bin ~ s(epitope.dist.b.ls) + southAfrica,
                                     family = binomial, data = train.dat,
                                   method = "NCV")
  fitted703.epitope <- predict(trainfit703.epitope, newdata = test.dat,
                               type = "response")
  
  testfit703.epitope[folds == j] <- fitted703.epitope
}

result <- ci.cvAUC(labels = amp703$tx_bin, predictions = testfit703.epitope, folds = folds)
fit703.epitope.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Hamming dist
testfit703.hdist <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.hdist <- mgcv::gam(tx_bin ~ s(hdist.zspace.sites.binding.all.ls) + southAfrica,
                                     family = binomial, data = train.dat,
                                 method = "NCV")
  fitted703.hdist <- predict(trainfit703.hdist, newdata = test.dat,
                             type = "response")
  
  testfit703.hdist[folds == j] <- fitted703.hdist
}

result <- ci.cvAUC(labels = amp703$tx_bin, predictions = testfit703.hdist, folds = folds)
fit703.hdist.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Raw IC80
testfit703.gmt80 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.gmt80 <- mgcv::gam(tx_bin ~ s(gmt80ls_num) + southAfrica,
                                     family = binomial, data = train.dat,
                                 method = "NCV")
  fitted703.gmt80 <- predict(trainfit703.gmt80, newdata = test.dat,
                             type = "response")
  
  testfit703.gmt80[folds == j] <- fitted703.gmt80
}

result <- ci.cvAUC(labels = amp703$tx_bin[!is.na(amp703$gmt80ls)],
                   predictions = testfit703.gmt80[!is.na(amp703$gmt80ls)],
                   folds = folds[!is.na(amp703$gmt80ls)])
fit703.gmt80.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r Objective1-704}
set.seed(300)
n <- nrow(amp704)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit704.parscore1.aucs <- numeric(5)

# Parscore 1
testfit704.parscore1 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.parscore1 <- mgcv::gam(tx_bin ~ s(parscore1.ls) + southAmerica,
                                 family = binomial, data = train.dat,
                                 method = "NCV")
  fitted704.parscore1 <- predict(trainfit704.parscore1, newdata = test.dat,
                                 type = "response")
  
  testfit704.parscore1[folds == j] <- fitted704.parscore1
}

result <- ci.cvAUC(labels = amp704$tx_bin, predictions = testfit704.parscore1, folds = folds)
fit704.parscore1.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Parscore 2
testfit704.parscore2 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.parscore2 <- mgcv::gam(tx_bin ~ s(parscore2.ls) + southAmerica,
                                     family = binomial, data = train.dat,
                                     method = "NCV")
  fitted704.parscore2 <- predict(trainfit704.parscore2, newdata = test.dat,
                                 type = "response")
  
  testfit704.parscore2[folds == j] <- fitted704.parscore2
}

result <- ci.cvAUC(labels = amp704$tx_bin, predictions = testfit704.parscore2, folds = folds)
fit704.parscore2.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Epitope dist
testfit704.epitope <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.epitope <- mgcv::gam(tx_bin ~ s(epitope.dist.b.ls) + southAmerica,
                                     family = binomial, data = train.dat,
                                   method = "NCV")
  fitted704.epitope <- predict(trainfit704.epitope, newdata = test.dat,
                               type = "response")
  
  testfit704.epitope[folds == j] <- fitted704.epitope
}

result <- ci.cvAUC(labels = amp704$tx_bin, predictions = testfit704.epitope, folds = folds)
fit704.epitope.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# Hamming dist
testfit704.hdist <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.hdist <- mgcv::gam(tx_bin ~ s(hdist.zspace.sites.binding.all.ls) + southAmerica,
                                     family = binomial, data = train.dat,
                                 method = "NCV")
  fitted704.hdist <- predict(trainfit704.hdist, newdata = test.dat,
                             type = "response")
  
  testfit704.hdist[folds == j] <- fitted704.hdist
}

result <- ci.cvAUC(labels = amp704$tx_bin, predictions = testfit704.hdist, folds = folds)
fit704.hdist.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)

# gmt 80
testfit704.gmt80 <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.gmt80 <- mgcv::gam(tx_bin ~ s(gmt80ls_num) + southAmerica,
                                     family = binomial, data = train.dat,
                                 method = "NCV")
  fitted704.gmt80 <- predict(trainfit704.gmt80, newdata = test.dat,
                             type = "response")
  
  testfit704.gmt80[folds == j] <- fitted704.gmt80
}

result <- ci.cvAUC(labels = amp704$tx_bin[!is.na(amp704$gmt80ls)],
                   predictions = testfit704.gmt80[!is.na(amp704$gmt80ls)],
                   folds = folds[!is.na(amp704$gmt80ls)])
fit704.gmt80.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r}
tab <- matrix(nrow = 5, ncol = 3)
tab[,1] <- c("PAR Score (logit)", "PAR Score (pred IC 80)", "Epitope Dist",
             "Hamming Dist", "Raw IC 80")
tab[1,2] <- paste0(fit703.parscore1.aucCI[1], " (", fit703.parscore1.aucCI[2],
                   ", ", fit703.parscore1.aucCI[3], ")")
tab[2,2] <- paste0(fit703.parscore2.aucCI[1], " (", fit703.parscore2.aucCI[2],
                   ", ", fit703.parscore2.aucCI[3], ")")
tab[3,2] <- paste0(fit703.epitope.aucCI[1], " (", fit703.epitope.aucCI[2],
                   ", ", fit703.epitope.aucCI[3], ")")
tab[4,2] <- paste0(fit703.hdist.aucCI[1], " (", fit703.hdist.aucCI[2],
                   ", ", fit703.hdist.aucCI[3], ")")
tab[5,2] <- paste0(fit703.gmt80.aucCI[1], " (", fit703.gmt80.aucCI[2],
                   ", ", fit703.gmt80.aucCI[3], ")")

tab[1,3] <- paste0(fit704.parscore1.aucCI[1], " (", fit704.parscore1.aucCI[2],
                   ", ", fit704.parscore1.aucCI[3], ")")
tab[2,3] <- paste0(fit704.parscore2.aucCI[1], " (", fit704.parscore2.aucCI[2],
                   ", ", fit704.parscore2.aucCI[3], ")")
tab[3,3] <- paste0(fit704.epitope.aucCI[1], " (", fit704.epitope.aucCI[2],
                   ", ", fit704.epitope.aucCI[3], ")")
tab[4,3] <- paste0(fit704.hdist.aucCI[1], " (", fit704.hdist.aucCI[2],
                   ", ", fit704.hdist.aucCI[3], ")")
tab[5,3] <- paste0(fit704.gmt80.aucCI[1], " (", fit704.gmt80.aucCI[2],
                   ", ", fit704.gmt80.aucCI[3], ")")

colnames(tab) <- c("Env AA Characteristic", "HVTN 703", "HVTN 704")


kable(tab)
```

## Ojective 2A: Prediction accuracy of generalized model including all Env AA features (excludng raw IC80)

```{r Objective2A-703}
set.seed(300)

n <- nrow(amp703)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit703.parscore1.aucs <- numeric(5)

# with gmt80
testfit703.all <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.all <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.b.ls) +
                                        s(hdist.zspace.sites.binding.all.ls) + southAfrica,
                                family = binomial, data = train.dat, method = "NCV")
  fitted703.all <- predict(trainfit703.all, newdata = test.dat, type = "response")
  
  testfit703.all[folds == j] <- fitted703.all
}

result <- ci.cvAUC(labels = amp703$tx_bin, predictions = testfit703.all, folds = folds)
fit703.all.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r Objective2A-704}
set.seed(300)

n <- nrow(amp704)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit704.parscore1.aucs <- numeric(5)


testfit704.all <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.all <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.b.ls) +
                                        s(hdist.zspace.sites.binding.all.ls) + southAmerica,
                                family = binomial, data = train.dat, method = "NCV")
  fitted704.all <- predict(trainfit704.all, newdata = test.dat, type = "response")
  
  testfit704.all[folds == j] <- fitted704.all
}

result <- ci.cvAUC(labels = amp704$tx_bin, predictions = testfit704.all, folds = folds)
fit704.all.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r Objective2A-results}
tab <- matrix(nrow = 1, ncol = 3)
tab[1,1] <- c("Full Model (no gmt80)")

tab[1,2] <- paste0(fit703.all.aucCI[1], " (", fit703.all.aucCI[2],
                   ", ", fit703.all.aucCI[3], ")")


tab[1,3] <- paste0(fit704.all.aucCI[1], " (", fit704.all.aucCI[2],
                   ", ", fit704.all.aucCI[3], ")")

colnames(tab) <- c("Env AA Characteristic", "HVTN 703", "HVTN 704")


kable(tab)
```

## Ojective 2B: Prediction accuracy of generalized model including all Env AA features (includng raw IC80)

```{r Objective2B-703}
set.seed(300)

n <- nrow(amp703)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit703.parscore1.aucs <- numeric(5)

# with gmt80
testfit703.all <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp703[folds != j,]
  test.dat <- amp703[folds == j,]
  
  trainfit703.all <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.b.ls) +
                                        s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica,
                                family = binomial, data = train.dat, method = "NCV")
  fitted703.all <- predict(trainfit703.all, newdata = test.dat, type = "response")
  
  testfit703.all[folds == j] <- fitted703.all
}

result <- ci.cvAUC(labels = amp703$tx_bin[!is.na(amp703$gmt80ls)], 
                   predictions = testfit703.all[!is.na(amp703$gmt80ls)],
                   folds = folds[!is.na(amp703$gmt80ls)])
fit703.allgmt80.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r Objective2B-704}
set.seed(300)

n <- nrow(amp704)
no.folds <- 5
folds <- sample(1:no.folds, n, replace = TRUE) 
fit704.parscore1.aucs <- numeric(5)

# with gmt80
testfit704.all <- numeric(n)
for(j in 1:no.folds) {
  train.dat <- amp704[folds != j,]
  test.dat <- amp704[folds == j,]
  
  trainfit704.all <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.b.ls) +
                                        s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica,
                                family = binomial, data = train.dat, method = "NCV")
  fitted704.all <- predict(trainfit704.all, newdata = test.dat, type = "response")
  
  testfit704.all[folds == j] <- fitted704.all
}

result <- ci.cvAUC(labels = amp704$tx_bin[!is.na(amp704$gmt80ls)], 
                   predictions = testfit704.all[!is.na(amp704$gmt80ls)], 
                   folds = folds[!is.na(amp704$gmt80ls)])
fit704.allgmt80.aucCI <- signif(c(result$cvAUC, result$ci), digits = 3)
```

```{r Objective2B-results}
tab <- matrix(nrow = 1, ncol = 3)
tab[1,1] <- c("Full Model (with gmt80)")

tab[1,2] <- paste0(fit703.allgmt80.aucCI[1], " (", fit703.allgmt80.aucCI[2],
                   ", ", fit703.allgmt80.aucCI[3], ")")


tab[1,3] <- paste0(fit704.allgmt80.aucCI[1], " (", fit704.allgmt80.aucCI[2],
                   ", ", fit704.allgmt80.aucCI[3], ")")

colnames(tab) <- c("Env AA Characteristic", "HVTN 703", "HVTN 704")


kable(tab)
```

## Objective 3A: Assessment of conditional dependence of each Env AA feature with TX Assignment (excluding raw IC80)

```{r Objective3A-703}
fit703.noparscore1 <- mgcv::gam(tx_bin ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                              family = binomial, data = amp703, method = "NCV")$fitted
fit703.noparscore2 <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                              family = binomial, data = amp703, method = "NCV")$fitted
fit703.noepitope <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) +
                                       s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                                       family = binomial, data = amp703, method = "NCV")$fitted
fit703.nohdist <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     southAfrica, 
                                     family = binomial, data = amp703, method = "NCV")$fitted

parscore1.pred703 <- mgcv::gam(parscore1.ls ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                                           family = gaussian, data = amp703, method = "NCV")$fitted

parscore2.pred703 <- mgcv::gam(parscore2.ls ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                                           family = gaussian, data = amp703, method = "NCV")$fitted

epitope.pred703 <- mgcv::gam(epitope.dist.c.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                                            family = gaussian, data = amp703, method = "NCV")$fitted

hdist.pred703 <- mgcv::gam(hdist.zspace.sites.binding.all.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(epitope.dist.c.ls) + southAfrica, 
                                            family = gaussian, data = amp703, method = "NCV")$fitted


parscore1.ccor.703 <- cor(amp703$parscore1.ls - parscore1.pred703,  
                          amp703$tx_bin - fit703.noparscore1)
parscore2.ccor.703 <- cor(amp703$parscore2.ls - parscore2.pred703,  
                          amp703$tx_bin - fit703.noparscore2)
epitope.ccor.703 <- cor(amp703$epitope.dist.c.ls - epitope.pred703,  
                        amp703$tx_bin - fit703.noepitope)
hdist.ccor.703 <- cor(amp703$hdist.zspace.sites.binding.all.ls - hdist.pred703,  
                        amp703$tx_bin - fit703.nohdist)

parscore1.ccor.703.p <- cor.test(amp703$parscore1.ls - parscore1.pred703,  
                                 amp703$tx_bin - fit703.noparscore1)$p.value
parscore2.ccor.703.p <- cor.test(amp703$parscore2.ls - parscore2.pred703,  
                                 amp703$tx_bin - fit703.noparscore2)$p.value
epitope.ccor.703.p <- cor.test(amp703$epitope.dist.c.ls - epitope.pred703,  
                               amp703$tx_bin - fit703.noepitope)$p.value
hdist.ccor.703.p <- cor.test(amp703$hdist.zspace.sites.binding.all.ls - hdist.pred703,  
                             amp703$tx_bin - fit703.nohdist)$p.value
```

```{r Objective3A-704}
fit704.noparscore1 <- mgcv::gam(tx_bin ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                              family = binomial, data = amp704, method = "NCV")$fitted
fit704.noparscore2 <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                              family = binomial, data = amp704, method = "NCV")$fitted
fit704.noepitope <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) +
                                       s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                                       family = binomial, data = amp704, method = "NCV")$fitted
fit704.nohdist <- mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     southAmerica, 
                                     family = binomial, data = amp704, method = "NCV")$fitted

parscore1.pred704 <- mgcv::gam(parscore1.ls ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                                           family = gaussian, data = amp704, method = "NCV")$fitted

parscore2.pred704 <- mgcv::gam(parscore2.ls ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                                           family = gaussian, data = amp704, method = "NCV")$fitted

epitope.pred704 <- mgcv::gam(epitope.dist.c.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                                            family = gaussian, data = amp704, method = "NCV")$fitted

hdist.pred704 <- mgcv::gam(hdist.zspace.sites.binding.all.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(epitope.dist.c.ls) + southAmerica, 
                                            family = gaussian, data = amp704, method = "NCV")$fitted


parscore1.ccor.704 <- cor(amp704$parscore1.ls - parscore1.pred704,  
                          amp704$tx_bin - fit704.noparscore1)
parscore2.ccor.704 <- cor(amp704$parscore2.ls - parscore2.pred704,  
                          amp704$tx_bin - fit704.noparscore2)
epitope.ccor.704 <- cor(amp704$epitope.dist.c.ls - epitope.pred704,  
                        amp704$tx_bin - fit704.noepitope)
hdist.ccor.704 <- cor(amp704$hdist.zspace.sites.binding.all.ls - hdist.pred704,  
                        amp704$tx_bin - fit704.nohdist)

parscore1.ccor.704.p <- cor.test(amp704$parscore1.ls - parscore1.pred704,  
                                 amp704$tx_bin - fit704.noparscore1)$p.value
parscore2.ccor.704.p <- cor.test(amp704$parscore2.ls - parscore2.pred704,  
                                 amp704$tx_bin - fit704.noparscore2)$p.value
epitope.ccor.704.p <- cor.test(amp704$epitope.dist.c.ls - epitope.pred704,  
                               amp704$tx_bin - fit704.noepitope)$p.value
hdist.ccor.704.p <- cor.test(amp704$hdist.zspace.sites.binding.all.ls - hdist.pred704,  
                             amp704$tx_bin - fit704.nohdist)$p.value
```


```{r Objective3A-results}
tab <- matrix(NA, ncol = 3, nrow = 4)

tab[,1] <- c("PAR Score (logit)", "PAR Score (pred IC 80)", "Epitope Dist",
             "Hamming Dist")

tab[1,2] <- paste0(signif(parscore1.ccor.703, 3), " (p = ",
                   signif(parscore1.ccor.703.p, 3),  ")")
tab[2,2] <- paste0(signif(parscore2.ccor.703, 3), " (p = ",
                   signif(parscore2.ccor.703.p, 3),  ")")
tab[3,2] <- paste0(signif(epitope.ccor.703, 3), " (p = ",
                   signif(epitope.ccor.703.p, 3),  ")")
tab[4,2] <- paste0(signif(hdist.ccor.703, 3), " (p = ",
                   signif(hdist.ccor.703.p, 3),  ")")

tab[1,3] <- paste0(signif(parscore1.ccor.704, 3), " (p = ",
                   signif(parscore1.ccor.704.p, 3),  ")")
tab[2,3] <- paste0(signif(parscore2.ccor.704, 3), " (p = ",
                   signif(parscore2.ccor.704.p, 3),  ")")
tab[3,3] <- paste0(signif(epitope.ccor.704, 3), " (p = ",
                   signif(epitope.ccor.704.p, 3),  ")")
tab[4,3] <- paste0(signif(hdist.ccor.704, 3), " (p = ",
                   signif(hdist.ccor.704.p, 3),  ")")

colnames(tab) <- c("Env AA Characteristic", "HVTN 703", "HVTN 704")

kable(tab)
```


## Objective 3B: Assessment of conditional dependence of each Env AA feature with TX Assignment (including raw IC80)

```{r Objective3B-703}
fit703.noparscore1 <- predict(mgcv::gam(tx_bin ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                              family = binomial, data = amp703, method = "NCV"),
                              newdata = amp703, type = "response")
fit703.noparscore2 <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                              family = binomial, data = amp703, method = "NCV"),
                              newdata = amp703, type = "response")
fit703.noepitope <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) +
                                       s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                                       family = binomial, data = amp703, method = "NCV"),
                            newdata = amp703, type = "response")
fit703.nohdist <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     s(gmt80ls_num) + southAfrica, 
                                     family = binomial, data = amp703, method = "NCV"),
                          newdata = amp703, type = "response")
fit703.nogmt80 <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     s(hdist.zspace.sites.binding.all.ls) + southAfrica, 
                                     family = binomial, data = amp703, method = "NCV"),
                          newdata = amp703, type = "response")



parscore1.pred703 <- predict(mgcv::gam(parscore1.ls ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                                           family = gaussian, data = amp703, method = "NCV"),
                             newdata = amp703)

parscore2.pred703 <- predict(mgcv::gam(parscore2.ls ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                                           family = gaussian, data = amp703, method = "NCV"),
                             newdata = amp703)

epitope.pred703 <- predict(mgcv::gam(epitope.dist.c.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAfrica, 
                                            family = gaussian, data = amp703, method = "NCV"),
                           newdata = amp703)

hdist.pred703 <- predict(mgcv::gam(hdist.zspace.sites.binding.all.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(gmt80ls_num) + s(epitope.dist.c.ls) + southAfrica, 
                                            family = gaussian, data = amp703, method = "NCV"),
                        newdata = amp703)

gmt80.pred703 <- predict(gam(gmt80ls_num ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + s(epitope.dist.c.ls) + southAfrica, 
                                            family = gaussian, data = amp703, method = "NCV"),
                         newdata = amp703)


parscore1.ccor.703 <- cor((amp703$parscore1.ls - parscore1.pred703)[!is.na(amp703$gmt80ls)],  
                          (amp703$tx_bin - fit703.noparscore1)[!is.na(amp703$gmt80ls)])
parscore2.ccor.703 <- cor((amp703$parscore2.ls - parscore2.pred703)[!is.na(amp703$gmt80ls)],  
                          (amp703$tx_bin - fit703.noparscore2)[!is.na(amp703$gmt80ls)])
epitope.ccor.703 <- cor((amp703$epitope.dist.c.ls - epitope.pred703)[!is.na(amp703$gmt80ls)],  
                        (amp703$tx_bin - fit703.noepitope)[!is.na(amp703$gmt80ls)])
hdist.ccor.703 <- cor((amp703$hdist.zspace.sites.binding.all.ls - hdist.pred703)[!is.na(amp703$gmt80ls)],  
                      (amp703$tx_bin - fit703.nohdist)[!is.na(amp703$gmt80ls)])
gmt80.ccor.703 <- cor((amp703$gmt80ls_num - gmt80.pred703)[!is.na(amp703$gmt80ls)],  
                    (amp703$tx_bin - fit703.nogmt80)[!is.na(amp703$gmt80ls)])

parscore1.ccor.703.p <- cor.test((amp703$parscore1.ls - parscore1.pred703)[!is.na(amp703$gmt80ls)],  
                                 (amp703$tx_bin - fit703.noparscore1)[!is.na(amp703$gmt80ls)])$p.value
parscore2.ccor.703.p <- cor.test((amp703$parscore2.ls - parscore2.pred703)[!is.na(amp703$gmt80ls)],  
                                 (amp703$tx_bin - fit703.noparscore2)[!is.na(amp703$gmt80ls)])$p.value
epitope.ccor.703.p <- cor.test((amp703$epitope.dist.c.ls - epitope.pred703)[!is.na(amp703$gmt80ls)],  
                               (amp703$tx_bin - fit703.noepitope)[!is.na(amp703$gmt80ls)])$p.value
hdist.ccor.703.p <- cor.test((amp703$hdist.zspace.sites.binding.all.ls - hdist.pred703)[!is.na(amp703$gmt80ls)],  
                             (amp703$tx_bin - fit703.nohdist)[!is.na(amp703$gmt80ls)])$p.value
gmt80.ccor.703.p <- cor.test((amp703$gmt80ls_num - gmt80.pred703)[!is.na(amp703$gmt80ls)],  
                           (amp703$tx_bin - fit703.nogmt80)[!is.na(amp703$gmt80ls)])$p.value
```

```{r Objective3B-704}
fit704.noparscore1 <- predict(mgcv::gam(tx_bin ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                              family = binomial, data = amp704, method = "NCV"),
                              newdata = amp704, type = "response")
fit704.noparscore2 <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                         s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                              family = binomial, data = amp704, method = "NCV"),
                              newdata = amp704, type = "response")
fit704.noepitope <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) +
                                       s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                                       family = binomial, data = amp704, method = "NCV"),
                            newdata = amp704, type = "response")

fit704.nohdist <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     s(gmt80ls_num) + southAmerica, 
                                     family = binomial, data = amp704, method = "NCV"),
                          newdata = amp704, type = "response")
fit704.nogmt80 <- predict(mgcv::gam(tx_bin ~ s(parscore1.ls) + s(parscore2.ls) + s(epitope.dist.c.ls) +
                                     s(hdist.zspace.sites.binding.all.ls) + southAmerica, 
                                     family = binomial, data = amp704, method = "NCV"),
                          newdata = amp704, type = "response")



parscore1.pred704 <- predict(mgcv::gam(parscore1.ls ~ s(parscore2.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                                           family = gaussian, data = amp704, method = "NCV"),
                             newdata = amp704)

parscore2.pred704 <- predict(mgcv::gam(parscore2.ls ~ s(parscore1.ls) + s(epitope.dist.c.ls) +
                                           s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                                           family = gaussian, data = amp704, method = "NCV"),
                             newdata = amp704)

epitope.pred704 <- predict(mgcv::gam(epitope.dist.c.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + s(gmt80ls_num) + southAmerica, 
                                            family = gaussian, data = amp704, method = "NCV"),
                           newdata = amp704)

hdist.pred704 <- predict(mgcv::gam(hdist.zspace.sites.binding.all.ls ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(gmt80ls_num) + s(epitope.dist.c.ls) + southAmerica, 
                                            family = gaussian, data = amp704, method = "NCV"),
                        newdata = amp704)

gmt80.pred704 <- predict(gam(gmt80ls_num ~ s(parscore1.ls) + s(parscore2.ls) +
                                            s(hdist.zspace.sites.binding.all.ls) + s(epitope.dist.c.ls) +                                                     southAmerica, 
                                            family = gaussian, data = amp704, method = "NCV"),
                         newdata = amp704)


parscore1.ccor.704 <- cor((amp704$parscore1.ls - parscore1.pred704)[!is.na(amp704$gmt80ls)],  
                          (amp704$tx_bin - fit704.noparscore1)[!is.na(amp704$gmt80ls)])
parscore2.ccor.704 <- cor((amp704$parscore2.ls - parscore2.pred704)[!is.na(amp704$gmt80ls)],  
                          (amp704$tx_bin - fit704.noparscore2)[!is.na(amp704$gmt80ls)])
epitope.ccor.704 <- cor((amp704$epitope.dist.c.ls - epitope.pred704)[!is.na(amp704$gmt80ls)],  
                        (amp704$tx_bin - fit704.noepitope)[!is.na(amp704$gmt80ls)])
hdist.ccor.704 <- cor((amp704$hdist.zspace.sites.binding.all.ls - hdist.pred704)[!is.na(amp704$gmt80ls)],  
                      (amp704$tx_bin - fit704.nohdist)[!is.na(amp704$gmt80ls)])
gmt80.ccor.704 <- cor((amp704$gmt80ls_num - gmt80.pred704)[!is.na(amp704$gmt80ls)],  
                    (amp704$tx_bin - fit704.nogmt80)[!is.na(amp704$gmt80ls)])

parscore1.ccor.704.p <- cor.test((amp704$parscore1.ls - parscore1.pred704)[!is.na(amp704$gmt80ls)],  
                                 (amp704$tx_bin - fit704.noparscore1)[!is.na(amp704$gmt80ls)])$p.value
parscore2.ccor.704.p <- cor.test((amp704$parscore2.ls - parscore2.pred704)[!is.na(amp704$gmt80ls)],  
                                 (amp704$tx_bin - fit704.noparscore2)[!is.na(amp704$gmt80ls)])$p.value
epitope.ccor.704.p <- cor.test((amp704$epitope.dist.c.ls - epitope.pred704)[!is.na(amp704$gmt80ls)],  
                               (amp704$tx_bin - fit704.noepitope)[!is.na(amp704$gmt80ls)])$p.value
hdist.ccor.704.p <- cor.test((amp704$hdist.zspace.sites.binding.all.ls - hdist.pred704)[!is.na(amp704$gmt80ls)],  
                             (amp704$tx_bin - fit704.nohdist)[!is.na(amp704$gmt80ls)])$p.value
gmt80.ccor.704.p <- cor.test((amp704$gmt80ls_num - gmt80.pred704)[!is.na(amp704$gmt80ls)],  
                           (amp704$tx_bin - fit704.nogmt80)[!is.na(amp704$gmt80ls)])$p.value
```


```{r Objective3B-results}
tab <- matrix(NA, ncol = 3, nrow = 5)

tab[,1] <- c("PAR Score (logit)", "PAR Score (pred IC 80)", "Epitope Dist",
             "Hamming Dist", "Raw IC80")

tab[1,2] <- paste0(signif(parscore1.ccor.703, 3), " (p = ",
                   signif(parscore1.ccor.703.p, 3),  ")")
tab[2,2] <- paste0(signif(parscore2.ccor.703, 3), " (p = ",
                   signif(parscore2.ccor.703.p, 3),  ")")
tab[3,2] <- paste0(signif(epitope.ccor.703, 3), " (p = ",
                   signif(epitope.ccor.703.p, 3),  ")")
tab[4,2] <- paste0(signif(hdist.ccor.703, 3), " (p = ",
                   signif(hdist.ccor.703.p, 3),  ")")
tab[5,2] <- paste0(signif(gmt80.ccor.703, 3), " (p = ",
                   signif(gmt80.ccor.703.p, 3),  ")")

tab[1,3] <- paste0(signif(parscore1.ccor.704, 3), " (p = ",
                   signif(parscore1.ccor.704.p, 3),  ")")
tab[2,3] <- paste0(signif(parscore2.ccor.704, 3), " (p = ",
                   signif(parscore2.ccor.704.p, 3),  ")")
tab[3,3] <- paste0(signif(epitope.ccor.704, 3), " (p = ",
                   signif(epitope.ccor.704.p, 3),  ")")
tab[4,3] <- paste0(signif(hdist.ccor.704, 3), " (p = ",
                   signif(hdist.ccor.704.p, 3),  ")")
tab[5,3] <- paste0(signif(gmt80.ccor.704, 3), " (p = ",
                   signif(gmt80.ccor.704.p, 3),  ")")

colnames(tab) <- c("Env AA Characteristic", "HVTN 703", "HVTN 704")

kable(tab)
```

```{r}
# marg.glm.703 <- glm(tx_bin ~ parscore1.ls + epitope.dist.c.ls + hdist.zspace.sites.binding.all.ls + 
#                epitope.dist.c.ls + southAfrica, 
#                family = binomial, data = amp703)
# 
# marg.glm.704 <- glm(tx_bin ~ parscore1.ls + epitope.dist.c.ls + hdist.zspace.sites.binding.all.ls + 
#                epitope.dist.c.ls + southAmerica, 
#                family = binomial, data = amp704)
# 
# glm.703 <- glm(tx_bin ~ parscore1.ls + epitope.dist.c.ls + hdist.zspace.sites.binding.all.ls + 
#                epitope.dist.c.ls + gmt80ls_num + southAfrica, 
#                family = binomial, data = amp703)
# 
# glm.704 <- glm(tx_bin ~ parscore1.ls + epitope.dist.c.ls + hdist.zspace.sites.binding.all.ls + 
#                epitope.dist.c.ls + gmt80ls_num + southAmerica, 
#                family = binomial, data = amp704)
```